---
title: "param2moment"
author: Tingting Zhan
format: 
  html:
    page-layout: full
    html-math-method: katex
toc: true
toc-location: left
toc-depth: 4
toc-title: ''
editor: source
knitr:
  opts_chunk: 
    collapse: true
    comment: "#>" 
bibliography: param2moment.bib
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette of package **`param2moment`** ([CRAN](https://CRAN.R-project.org/package=param2moment), [Github](https://github.com/tingtingzhan/param2moment)) documents ...

## Prerequisite

New features are first implemented on [Github](https://github.com/tingtingzhan/param2moment).

```{r}
#| warning: false
#| eval: false
remotes::install_github('tingtingzhan/TukeyGH77')
remotes::install_github('tingtingzhan/param2moment')
```

And eventually make their way to [`CRAN`](https://CRAN.R-project.org/package=param2moment).

```{r}
#| warning: false
#| eval: false
utils::install.packages('TukeyGH77')
utils::install.packages('param2moment')
```

## Note to Users

Examples in this vignette require that the `search` path has

```{r}
#| message: false
library(param2moment)
library(ggplot2)
library(geomtextpath)
library(TukeyGH77)
library(sn)
```

```{r}
#| echo: false
op = par(no.readonly = TRUE)
theme_bw() |>
  theme_set()
```

# Introduction

# Moments

For random variable $X$,

## Raw Moments

| Slot    | Description         |
|---------|---------------------|
| `@raw1` | $\mu = \text{E}(X)$ |
| `@raw2` | $\text{E}(X^2)$     |
| `@raw3` | $\text{E}(X^3)$     |
| `@raw4` | $\text{E}(X^4)$     |

## Central Moments

According to [Binomial theorem](https://en.wikipedia.org/wiki/Binomial_theorem),

| Slot | [Central Moment](https://en.wikipedia.org/wiki/Central_moment) | Value | Derivation |
|------------------|------------------|------------------|------------------|
|  | $0 = \text{E}(X-\mu)$ |  |  |
| `@central2` | $\sigma^2 = \text{E}(X-\mu)^2$ | $\text{E}(X^2) - \mu^2$ | $\text{E}(X^2) - 2\mu\text{E}(X) + \mu^2$ |
| `@central3` | $\text{E}(X-\mu)^3$ | $\text{E}(X^3) - 3\mu\text{E}(X^2) + 2\mu^3$ | $\text{E}(X^3) - 3\mu\text{E}(X^2) + 3\mu^2\text{E}(X) - \mu^3$ |
| `@central4` | $\text{E}(X-\mu)^4$ | $\text{E}(X^4) - 4\mu\text{E}(X^3) + 6\mu^2\text{E}(X^2) - 3\mu^4$ | $\text{E}(X^4) - 4\mu\text{E}(X^3) + 6\mu^2 \text{E}(X^2) - 4\mu^3\text{E}(X) + \mu^4$ |

## Standardized Moments

| Slot | [Standardized Moment](https://en.wikipedia.org/wiki/Standardized_moment) | Description |
|------------------------|------------------------|------------------------|
|  | $0 = \text{E}(X-\mu)/\sigma$ |  |
|  | $1 = \text{E}(X-\mu)^2/\sigma^2$ |  |
| `@standardized3` | $\text{E}(X-\mu)^3/\sigma^3$ | [skewness](https://en.wikipedia.org/wiki/Skewness) |
| `@standardized4` | $\text{E}(X-\mu)^4/\sigma^4$ | [kurtosis](https://en.wikipedia.org/wiki/Kurtosis) |

## Location-Scale Transformation

For $Y = \text{location} + \text{scale} \cdot X$,

$$
\begin{cases}
\mu_Y  & = \text{location} + \text{scale}\cdot \mu_X\\
\sigma_Y & = \text{scale}\cdot \sigma_X \\
\text{skewness}_Y & = \text{skewness}_X\\
\text{kurtosis}_Y & = \text{kurtosis}_X
\end{cases}
$$

# Conversion of Parameter and Moment

## Normal Distribution

```{r}
moment_norm(mean = 1.2, sd = .7)
```

## [Skew-Normal](https://en.wikipedia.org/wiki/Skew_normal_distribution) Distribution

```{r}
moment_sn(xi = 2, omega = 1.3, alpha = 3)
```

```{r}
moment2sn(skewness = .3)
```

## Skew-$t$ Distribution

```{r}
moment_st(xi = 2, omega = 1.3, alpha = 3, nu = 6)
```

```{r}
moment2st(skewness = .2, kurtosis = .3)
```

```{r}
moment2st(mean = 1, kurtosis = .3, alpha = 0)
```

## Tukey's $gh$ Distribution

### Tukey's $g$-distribution

```{r}
moment_GH(A = 3, B = 1.5, g = 0, h = .05)
```

```{r}
moment2GH(mean = 1, kurtosis = .3, g = 0)
```

### Tukey's $h$-distribution

```{r}
moment_GH(A = 3, B = 1.5, g = .7, h = 0)
```

```{r}
moment2GH(skewness = .2, h = 0)
```

### Tukey's $g$-and-$h$ Distribution

```{r}
moment_GH(A = 3, B = 1.5, g = .7, h = .05)
```

```{r}
moment2GH(skewness = .2, kurtosis = .3)
```

## [Gamma Distribution](https://en.wikipedia.org/wiki/Gamma_distribution)

# Batch Visualization

```{r}
skw = c(.2, .5, .8)
krt = c(.5, 1, 1.5)
```

## Kurtosis only

@Student:08-$t$ vs. Tukey's $h$

-   R function `sn::dst` [@R_sn]   [scale $\omega$, degree-of-freedom $\nu$]{style="background-color:yellow;"}

-   $\mu = 0$, $\sigma = 1$, [Kurtosis]{style="font-weight:bold;background-color:yellow;"}=[.5]{style="color:red;font-weight:bold;"}, [1]{style="color:green;font-weight:bold;"}, or [1.5]{style="color:blue;font-weight:bold;"}

```{r}
#| fig-height: 3
#| fig-width: 10
#| code-fold: true
#| code-summary: '[R] Student-t and Tukey-h'
p1 = moment2param(distname = 'st', MoreArgs = list(alpha = 0), kurtosis = krt) + 
  xlim(-1.5, 1.5) + labs(y = latex2exp::TeX('Student-$t$'))
p2 = moment2param(distname = 'GH', MoreArgs = list(g = 0), kurtosis = krt) + 
  xlim(-1.5, 1.5) + labs(y = latex2exp::TeX('Tukey\'s $h$'))
p1 + p2
```

## Skewness only

Skew-Normal vs. Tukey's $g$

-   R function `sn::dsn`   [location $\xi$, scale $\omega$, slant $\alpha$]{style="background-color:yellow;"}

-   $\mu = 0$, $\sigma = 1$, [Skewness]{style="font-weight:bold;background-color:yellow;"}=[.2]{style="color:red;font-weight:bold;"}, [.5]{style="color:green;font-weight:bold;"}, or [.8]{style="color:blue;font-weight:bold;"}

```{r}
#| fig-height: 3
#| fig-width: 10
#| code-fold: true
#| code-summary: '[R] Skew-Normal and Tukey-g'
p1 = moment2param(dist = 'sn', skewness = skw, hjust = .42) + xlim(-2, 2.5) + labs(y = 'Skew-Normal')
p2 = moment2param(dist = 'GH', MoreArgs = list(h = 0), skewness = skw, hjust = .42) + 
  xlim(-2, 2.5) + labs(y = latex2exp::TeX('Tukey\'s $g$'))
p1 + p2
```

## Skewness & Kurtosis

Skew-$t$ vs. Tukey's $g$-&-$h$

-   R function `sn::dst`   [location $\xi$, scale $\omega$, slant $\alpha$, degree-of-freedom $\nu$]{style="background-color:yellow;"}

-   $\mu = 0$, $\sigma = 1$, [(Skewness, Kurtosis)]{style="font-weight:bold;background-color:yellow;"}=[(.2,.5)]{style="color:red;font-weight:bold;"}, [(.5,1)]{style="color:green;font-weight:bold;"}, or [(.8,1.5)]{style="color:blue;font-weight:bold;"}

```{r}
#| fig-height: 3
#| fig-width: 10
#| code-fold: true
#| code-summary: '[R] Skew-t and Tukey-gh'
p1 = moment2param(distname = 'st', skewness = skw, kurtosis = krt, hjust = .35) +
  xlim(-2, 2.5) + labs(y = latex2exp::TeX('Skew-$t$'))
p2 = moment2param(distname = 'GH', skewness = skw, kurtosis = krt, hjust = .35) +
  xlim(-2, 2.5) + labs(y = latex2exp::TeX('Tukey\'s $g$-&-$h$'))
p1 + p2
```

```{r}
#| fig-height: 3
#| fig-width: 5
#| code-fold: true
#| code-summary: '[R] Gamma'
moment2param(distname = 'gamma', skewness = skw, kurtosis = 3, hjust = .4) +
  xlim(0, 4) + labs(y = latex2exp::TeX('$gamma$'))
```


# Appendix 

## Terms and Abbreviations

| Term / Abbreviation | Description |
|------------------------------------|------------------------------------|
| [`|>`](https://search.r-project.org/R/refmans/base/html/pipeOp.html) | Forward pipe operator introduced in `R` 4.1.0 |
| [`all.equal.numeric`](https://search.r-project.org/R/refmans/base/html/all.equal.html) | Test of near equality |
| `CRAN`, `R` | [The Comprehensive R Archive Network](https://cran.r-project.org) |
| [`curve`](https://search.r-project.org/R/refmans/graphics/html/curve.html) | Function plots |
| [`.lm.fit`](https://search.r-project.org/R/refmans/stats/html/lmfit.html) | Least squares regression, the internal workhorse |
| [`mad`](https://search.r-project.org/R/refmans/stats/html/mad.html) | Median Absolute Deviation |
| [`median`](https://search.r-project.org/R/refmans/stats/html/median.html) | Median |
| [`dnorm`, `pnorm`, `qnorm`, `rnorm`](https://search.r-project.org/R/refmans/stats/html/Normal.html) | Normal Distribution |
| [`optim`](https://search.r-project.org/R/refmans/stats/html/optim.html), [`optimize`](https://search.r-project.org/R/refmans/stats/html/optimize.html) | Optimization |
| [`sd`](https://search.r-project.org/R/refmans/stats/html/sd.html) | Standard Deviation |
| [`search`](https://search.r-project.org/R/refmans/base/html/search.html) | Search path |
| [`seed`](https://search.r-project.org/R/refmans/base/html/Random.html) | Random number generation seed |
| [`vuniroot`](https://search.r-project.org/CRAN/refmans/rstpm2/html/vuniroot.html) | Vectorised one-dimensional root-finding, in package [`rstpm2`](https://CRAN.R-project.org/package=rstpm2) |


